Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-11 manuscript/manuscript.tex
editing Ne section

We observed that in the longest study  we analyzed \parencite{Barghi2019-qy},
some genome-wide temporal covariances become negative at future timepoints (see
the first two rows in Figure \ref{fig:figure-1} A). This shows that alleles
that were on average going up initially are later going down in frequency, i.e.
that the average direction of selection experienced by alleles has flipped.
This must reflect either a change in the environment or the genetic background,
due to epistatic relationships among alleles altered by frequency changes or
recombination breaking up selective alleles.  Such reversals of selective
dynamics could be occurring at other timepoints but the signal of a change in
the direction of selection at particular loci may be washed out when we
calculate our genome-wide average temporal covariances. To address this
limitation, we calculated the distribution of the temporal covariances over
100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
pooling across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance estimate of each genomic window will be noisy, due to sampling and
genetic drift, and the neutral distribution of the covariance is complicated
due to linkage disequilibria, which can occur over long physical distances in
E\&R and selection studies \parencite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}. To
address this, we have developed a permutation-based procedure that constructs
an empirical null distribution by randomly flipping the signs of the allele
frequency changes per-genomic window. This destroys the systematic covariances
created by linked selection and creates a sampling distribution of the
covariances spuriously created by neutral genetic drift while preserving the
complex dependencies between adjacent loci created by linkage disequilibrium.
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift alone as they
include the effect of selection on the allele frequency change within a
time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

>>>>

We observed that in the longest study  we analyzed \parencite{Barghi2019-qy},
some genome-wide temporal covariances become negative at future timepoints (see
the first two rows in Figure \ref{fig:figure-1} A). This shows that alleles
that were on average going up initially are later going down in frequency, i.e.
that the average direction of selection experienced by alleles has flipped.
This must reflect either a change in the environment or the genetic background,
due to epistatic relationships among alleles altered by frequency changes or
recombination breaking up selective alleles.  Such reversals of selective
dynamics could be occurring at other timepoints but the signal of a change in
the direction of selection at particular loci may be washed out when we
calculate our genome-wide average temporal covariances. To address this
limitation, we calculated the distribution of the temporal covariances over
100kb windowed covariances (Figure \ref{fig:figure-3} shows these distributions
pooling across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance estimate of each genomic window will be noisy, due to sampling and
genetic drift, and the neutral distribution of the covariance is complicated
due to linkage disequilibria, which can occur over long physical distances in
E\&R and selection studies \parencite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}. To
address this, we have developed a permutation-based procedure that constructs
an empirical null distribution by randomly flipping the \gc{sign }of the allele
frequency changes \gc{in each} genomic window \gc{(i.e. a single random sign
flip is applied to all loci in a window)}. This destroys the systematic covariances
created by linked selection and creates a sampling distribution of the
covariances spuriously created by neutral genetic drift while preserving the
complex dependencies between adjacent loci created by linkage disequilibrium.
This empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift alone as they
include the effect of selection on the allele frequency change within a
time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) \gc{that there are an empirical excess of windows
  with positive
  covariances between close timepoints compared to the null
  distribution (a heavier right tail), and that this then shifts to an excess of windows
  with negative covariances between more distant
timepoints (a heavier left tail)}
%that windowed temporal covariances between close
t%imepoints are skewed positive (a heavy right tail), while between more distant
%timepoints these windowed temporal covariances tend to shift to become more
%negative (a heavy left tail).
We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.
